Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-02 document/ms.tex
some discussion of simulations

To confirm that the probabilistic framework developed in the previous sections
provides correct inferences about the population of exoplanets given noisy
data, we'll start by generating synthetic catalogs and running our full
inference pipeline on those data.
In this section, we also demonstrate the key benefits of our method over the
standard intuitive inverse-detection-efficiency procedure.

\paragraph{Inverse-detection-efficiency}
Authors in the exoplanet literature (for example, \citealt{howard, dressing,
petigura}) have built population level models using this procedure.
The basic idea is to make a histogram of planet candidates in the model where
each sample has a weight proportional to the inverse detection efficiency.
This method makes intuitive sense and, in the limit of infinitely precise
measurements, it is not wrong.
Even in this limit, this estimator of the occurrence distributions will have
large variance; especially as the detection efficiency becomes low.
In the presence of observational uncertainties, the inferred distributions
will be biased.

The likelihood function in \eq{true-like} provides a \emph{probability
distribution over catalogs} conditioned on the parameters of the rate
function.
Therefore, it is also the description of a \emph{generative process} for
synthetic catalogs.
To demonstrate the effectiveness of our model, we generate synthetic catalogs
from known distributions in period and radius, and test the performance of our
method on the resulting data.
For comparison, we also implement the one-over-efficiency weighting procedure
(see, for example, \citealt{howard, dressing, petigura}) and apply it to the
same datasets.

\paragraph{The generative process}
The variable rate Poisson likelihood function in \eq{true-like} can be sampled
generally for arbitrary forms of the rate function using a procedure called
``thinning'' (\citealt{poisson}).
In this \paper, however, we will only consider binned representations of
\rate, greatly simplifying the generative process.
Conditioned on a set of rate function parameters and a given completeness
function (here we'll use the injections from \citealt{petigura}), we compute
the observable rate grid from \eq{obs-rate}.
Then in each bin, we sample $J$ candidates uniformly in period and
radius (as defined by the bin edges) where $J$ is a Poisson random variable
sampled with the rate given by the observable rate in this bin.
\dfmtodo{Make this description not be the most confusing thing ever.}
Then, we add 30\% measurement uncertainties to the radius measurements where
30\% was chosen to match to properties of the real catalog
(\citealt{petigura}).

>>>>

In order to get a feeling for the constraints provided by our method and to
explore any biases introduced by ignoring the observational uncertainties, we
start by ``observing'' a few synthetic catalogs from qualitatively different
known occurence rate functions.
For each of these simulations, we'll take the completeness function computed
by \citet{petigura} as given.
In general, \eq{poisson-like} can be sampled using a procedure called thinning
(\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
constant rate function evaluated on a fine grid in (log-) period and radius.
For this discrete function, the generative procedure is simply:
\begin{enumerate}
{\item loop over each grid cell $i$,}
{\item draw Poisson random integer $K_i\sim\mathrm{Poissson}(\obs{\rate}_i)$
with the observable rate evaluated in the cell, and}
{\item distribute $K_i$ catalog entries in the cell randomly.}
\end{enumerate}
We then choose fractional observational uncertainties on the radii from the
\citet{petigura} catalog and apply them to the catalog assuming that they are
Gaussian.
